---
title: "CSC3002F - Operating Systems II"
subtitle: "Scheduling Assignment"
author: "Connor Morrison (MRRCON006)"
date: today
date-format: "Do MMMM YYYY"
format: 
  pdf:
    toc: true
    fig-width: 8
    fig-height: 5
    knitr:
      opts_chunk: 
        dev: "png"
        dpi: 300
  
---

```{r, include=FALSE}
pkg_vec <- c("ggplot2","ggthemes","ggh4x","broom","kableExtra","tidyr","tidyverse","emmeans","effectsize","car","dplyr")
for (x in pkg_vec) {
  if (!requireNamespace(x, quietly = TRUE)) {
    install.packages(x, dependencies = TRUE)
  }
  library(x, character.only = TRUE)
}

```

```{r, include=FALSE}
df = read_csv("simulation_results.csv")
df$total_patrons = factor(df$total_patrons)
df$schedalg = factor(df$schedalg, levels = c(0, 1, 2, 3), labels = c("FCFS", "SJF", "Priority", "MLFQ"))
df$drink = factor(df$drink)
df$priority = factor(df$priority)
df$queue_level = factor(df$queue_level)
```

# Introduction

Allegra the barman is an abstraction of the low-level workings of a CPU. The study of CPU scheduling algorithms can be likened much to the analogy of a barman serving drinks at a bar, where the overall goal is to most efficiently and effectively serve all customers such that the majority of customers will be satisfied with their experience.

The goal of this assignment is to determine the CPU scheduling algorithm that, when applied to the bar setting, maximises satisfaction for the most customers.

# Methods

To generate the data


# Results

```{r,echo=FALSE}

averages <- df %>% group_by(total_patrons,schedalg, patron) %>% 
  summarise(avg_waiting = mean(waiting_time),std_waiting = var(waiting_time), avg_turnaround = 
            mean(turnaround_time), std_turnaround = var(turnaround_time), avg_response = mean(response_time), 
            std_response = var(response_time), num_orders = n(), .groups = 'drop')

ggplot(averages,aes(x=avg_response,fill=schedalg)) + geom_density(alpha = 0.5) + facet_wrap(~total_patrons, ncol = 1, scales = "free") + 
  facetted_pos_scales(
    x = list(
      total_patrons == 5 ~ scale_x_continuous(limits = c(0, 50)),
      total_patrons == 20 ~ scale_x_continuous(limits = c(0, 500)),
      total_patrons == 50 ~ scale_x_continuous(limits = c(0, 2500))
    )) + 
  labs(title = "Distribution of Average Response Time by Workload",
       x = "Average Response Time (ms)",
       y = "Density")

ggplot(averages,aes(x=avg_waiting,fill=schedalg)) + geom_density(alpha = 0.5) + facet_wrap(~total_patrons, ncol = 1, scales = "free") + 
  facetted_pos_scales(
    x = list(
      total_patrons == 5 ~ scale_x_continuous(limits = c(0, 50)),
      total_patrons == 20 ~ scale_x_continuous(limits = c(0, 500)),
      total_patrons == 50 ~ scale_x_continuous(limits = c(0, 2500))
    )) + 
  labs(title = "Distribution of Average Waiting Time by Workload",
       x = "Average Waiting Time (ms)",
       y = "Density")

ggplot(averages,aes(x=avg_turnaround,fill=schedalg)) + geom_density(alpha = 0.5) + facet_wrap(~total_patrons, ncol = 1, scales = "free")  + 
  facetted_pos_scales(
    x = list(
      total_patrons == 5 ~ scale_x_continuous(limits = c(30, 100)),
      total_patrons == 20 ~ scale_x_continuous(limits = c(0, 500)),
      total_patrons == 50 ~ scale_x_continuous(limits = c(0, 2500))
    )) + 
  labs(title = "Distribution of Average Turnaround Time by Workload",
       x = "Average Turnaround Time (ms)",
       y = "Density")


stats_df <- averages %>% group_by(total_patrons, schedalg) %>%
  summarise(mean_val = mean(avg_waiting),
            median_val = median(avg_waiting), 
            .groups = 'drop')

ggplot(averages, aes(x = avg_waiting, fill = schedalg)) + 
  geom_density(alpha = 0.5) +
  # Add Mean lines
  geom_vline(data = stats_df, aes(xintercept = mean_val, color = schedalg), 
             linetype = "solid", linewidth = 0.5) +
  # Add Median lines
  geom_vline(data = stats_df, aes(xintercept = median_val, color = schedalg), 
             linetype = "dashed", linewidth = 0.5) +
  facet_wrap(~total_patrons, ncol = 1, scales = "free") + 
  facetted_pos_scales(
    x = list(
      total_patrons == 5 ~ scale_x_continuous(limits = c(0, 50)),
      total_patrons == 20 ~ scale_x_continuous(limits = c(0, 500)),
      total_patrons == 50 ~ scale_x_continuous(limits = c(0, 2500))
    )
  ) +
  labs(title = "Waiting Time Distributions with Mean (Solid) and Median (Dashed)",
       x = "Average Waiting Time (ms)",
       y = "Density")


summary_plot_df <- stats_df %>% pivot_longer(cols = c(mean_val, median_val), names_to = "Metric", values_to = "Value")
ggplot(summary_plot_df, aes(x = Value, y = schedalg, color = Metric)) +
  geom_line(aes(group = schedalg), color = "grey") + # Connects Mean and Median with a line
  geom_point(size = 3) +
  facet_wrap(~total_patrons, ncol = 1, scales = "free_x") +
  labs(title = "Mean vs Median Wait Time Gap by Algorithm",
       x = "Waiting Time (ms)",
       y = "Algorithm")

#uses pivoted summary statistics to ensure that the mean and median were based on the exact same runs
```

```{r}
ggplot(df, aes(x=schedalg, y=waiting_time, fill=schedalg)) +
  geom_boxplot() +
  facet_wrap(~total_patrons, scales = "free") +
  labs(title="Waiting Time Distribution by Algorithm")

# 1. Fit the model
fit <- lm(waiting_time ~ schedalg + total_patrons + drink, data = df)

# 2. View the significance (p-values) and magnitudes (coefficients)
summary(fit)

# 3. Check for Fairness/Starvation by plotting residuals
plot(fit, which = 1)
```
```{r}
# Compare means
model_speed <- lm(avg_waiting ~ schedalg, data = averages)
summary(model_speed)
plot(model_speed)
Anova(model_speed)
```

# Conclusions





























